Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-06-06 PAPER.tex
Wed Jun  6 11:44:31 EDT 2018

% Math shortcuts from James
\newcommand{\invC}{\ensuremath{\C^{-1}}}
\newcommand{\dC}[1]{\ensuremath{\C_{,#1}}}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\half}{\ensuremath{\frac{1}{2}}}
\newcommand{\PDeriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}

@@

% JEA - The discussion of dimensions here is confusing ... "a baseline" is mentioned, but it's not clear if the redundant baseline averaging has already happened or what.  I believe we are only considering covariance between frequencies, but this isn't explicitly said, and the matrix equations don't make sense unless all the dimensions are N_f N_t.  
%We begin with our 
The measured visibilities for a single baseline in Jy are arranged as a data vector, $\textbf{x}$. It has length $N_{t} N_{f}$,
% JEA - not sure this helps
%, but in practice we manipulate it as an array with dimensions ($N_{f}, N_{t}$), 
where $N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. The covariance of the data is given by 
\begin{equation}
\textbf{C} \equiv \langle\textbf{xx}^{\dagger}\rangle = \textbf{S} + \textbf{U}
\end{equation}
where we assume the covariance can be written as a sum of the desired cosmological signal $\textbf{S}$ and other terms $\textbf{U}$.

We are interested in estimating the three-dimensional power spectrum of the EoR.  
Visibilities are measurements of the Fourier transform of the sky along two spatial dimensions (using the flat-sky approximation), and since we are interested in three-dimensional Fourier modes we only need to take one Fourier transform of our visibilities along the line-of-sight dimension.  We consider band powers $p_\alpha$ of the power spectrum of $\textbf{x}$ over some range in cosmological $\mathbf{k}$, where $\alpha$ indexes a waveband in $k_{\parallel}$ (a cosmological wavenumber $k_{\parallel}$ is the Fourier dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}), which is a good approximation for the short baselines that PAPER analyzes).  The fundamental dependence of the covariance on the power spectrum band powers $p_\alpha$ is encoded as 
% Should distinguish slightly more clearly between Q and dC/dp_alpha
\begin{equation}
%\textbf{S} = p_{\alpha} 
\frac{\partial\textbf{C}}{\partial p_{\alpha}} \equiv \textbf{Q}^{\alpha}
\end{equation}
which has dimensions ($N_{f}$,$N_{f}$).
In practice, this matrix also implements a transform of the frequency-domain visibilities to $\mathbf{k}$-space, taking into account cosmological scalings, and converting the visibilities from Jy to Kelvin.

The optimal quadratic estimator prescription is then to compute
\begin{equation}
\label{eq:OQE}
\widehat{p}_{\alpha}  =  \half ({\F}^{\alpha \beta})^{-1} (\widehat{q}_{\beta} + \widehat{b}_{\beta} )
\end{equation}
where $\F$ is the Fisher matrix and
\begin{equation}
\label{eq:OQEQuadratic}
\hat{q}_\beta =  \textbf{x}^\dagger \invC \PDeriv{\C}{p_\beta}  \invC \textbf{x} 
\end{equation}
and
\begin{equation}
\label{eq:OQELinear}
\widehat{b}_{\beta} = \Tr\left[ (\C - \mathbf{S}) \invC \PDeriv{\C}{p_\beta} \invC \right].
\end{equation}
This estimator is the minimum variance (smallest error bar) estimate of the power spectrum \citep{tegmark_et_al1997a,bond_et_al1998} and is unbiased if the ensemble average of the estimator is equal to its true value
\begin{equation}
\langle \widehat{p}_{\alpha} \rangle = p_\alpha.
\end{equation}

@@

%We do this when forming the un-normalized power spectrum estimate $\widehat{q}_{\alpha}$:
It is also possible to define a more generalized version of the optimal quadratic estimator which may still produce an unbiased estimate, where Equation \eqref{eq:OQEQuadratic} is replaced by
\begin{equation}
\label{eq:qhat}
\widehat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}.
\end{equation}
where $\textbf{R}$ is a weighting matrix chosen by the data analyst.
%with dimensions ($N_{f}$,$N_{f}$) --- as 
%an example, inverse covariance weighting (the optimal form of QE) would set $\textbf{R} \equiv \textbf{C}^{-1}$ and a uniform-weighted case would use $
%\textbf{R} \equiv \textbf{I}$, the identity matrix.
With an appropriate normalization 
%We normalize our power spectrum estimates using the matrix $\textbf{M}$:
\begin{equation}
\label{eq:phat}
\widehat{\textbf{P}} = \textbf{M}\widehat{\textbf{q}},
\end{equation}
$\widehat{\textbf{P}}$ is the estimate of the true power spectrum $\textbf{P}$. The data analyst has a choice for $
\textbf{M}$. For simplicity in this section we choose $\textbf{M}$ to be diagonal, although we explore other cases for the analysis 
of PAPER-64 data as explained in Section \ref{sec:Bias}, subject to the constraint that each element of $\widehat{\textbf{P}}$ is a weighted sum of the elements of $\textbf{P}$ with weights that sum to unity.
% JEA - is one really free to choose M and R independently?

>>>>

% Math shortcuts from James
\newcommand{\invC}{\ensuremath{\C^{-1}}}
\newcommand{\dC}[1]{\ensuremath{\C_{,#1}}}
\DeclareMathOperator{\Tr}{tr}
\newcommand{\half}{\ensuremath{\frac{1}{2}}}
\newcommand{\PDeriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}

@@

% JEA - The discussion of dimensions here is confusing ... "a baseline" is mentioned, but it's not clear if the redundant baseline averaging has already happened or what.  I believe we are only considering covariance between frequencies, but this isn't explicitly said, and the matrix equations don't make sense unless all the dimensions are N_f N_t.  
%We begin with our 
The measured visibilities for a single baseline in Jy are arranged as a data vector, $\textbf{x}$. It has length $N_{t} N_{f}$,
% JEA - not sure this helps
%, but in practice we manipulate it as an array with dimensions ($N_{f}, N_{t}$), 
where $N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. The covariance of the data is given by 
\begin{equation}
\textbf{C} \equiv \langle\textbf{xx}^{\dagger}\rangle = \textbf{S} + \textbf{U}
\end{equation}
where the average over an ensemble of data realizations produces the true covariance, and we further assume it may be written as a sum of the desired cosmological signal $\textbf{S}$ and other terms $\textbf{U}$.

We are interested in estimating the three-dimensional power spectrum of the EoR.  
Visibilities are measurements of the Fourier transform of the sky along two spatial dimensions (using the flat-sky approximation), and since we are interested in three-dimensional Fourier modes we only need to take one Fourier transform of our visibilities along the line-of-sight dimension.  We consider band powers $p_\alpha$ of the power spectrum of $\textbf{x}$ over some range in cosmological $\mathbf{k}$, where $\alpha$ indexes a waveband in $k_{\parallel}$ (a cosmological wavenumber $k_{\parallel}$ is the Fourier dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}), which is a good approximation for the short baselines that PAPER analyzes).  The fundamental dependence of the covariance on the power spectrum band powers $p_\alpha$ is encoded as 
% JEA - Should distinguish slightly more clearly between Q and dC/dp_alpha
\begin{equation}
\textbf{S} = p_{\alpha} \frac{\partial\textbf{C}}{\partial p_{\alpha}}
%\frac{\partial\textbf{C}}{\partial p_{\alpha}} \equiv \textbf{Q}^{\alpha}
\end{equation}
where we use the Einstein summation convention.
% JEA - ah, the dimensions again
% which has dimensions ($N_{f}$,$N_{f}$).

The optimal quadratic estimator prescription is then to compute
\begin{equation}
\label{eq:OQE}
\widehat{p}_{\alpha}  =  ({\F^{-1}})_\alpha^{\beta} (\widehat{q}_{\beta} - \widehat{b}_{\beta} )
\end{equation}
where $\F$ is the Fisher matrix 
\begin{equation}
F^{\alpha \beta} \equiv \frac{1}{2} \textrm{tr} \left( \C^{-1} \frac{\partial\textbf{C}}{\partial p_{\alpha}} \C^{-1} \frac{\partial\textbf{C}}{\partial p_{\beta}} \right)
\end{equation}
and
\begin{equation}
\label{eq:OQEQuadratic}
\hat{q}_\alpha =  \half \textbf{x}^\dagger \invC \PDeriv{\C}{p^\alpha}  \invC \textbf{x} 
\end{equation}
and
\begin{equation}
\label{eq:OQELinear}
\widehat{b}_{\alpha} = \half \Tr\left[ (\C - \mathbf{S}) \invC \PDeriv{\C}{p^\alpha} \invC \right].
\end{equation}
This estimator is the minimum variance (smallest error bar) estimate of the power spectrum  subject to the constraint that it is also unbiased; that is, the ensemble average of the estimator is equal to its true value
\begin{equation}
\langle \widehat{p}_{\alpha} \rangle = p_\alpha.
\end{equation}
\citep{tegmark_et_al1997a,bond_et_al1998}

@@

%We do this when forming the un-normalized power spectrum estimate $\widehat{q}_{\alpha}$:
Motivated by the desire to retain the desirable behavior of suppressing contributions of $\mathbf{U}$ to estimates of the EoR power spectrum, we note that is possible to define a more generalized version of the quadratic estimator 
%which may still produce an unbiased estimate, 
where Equation \eqref{eq:OQEQuadratic} is replaced by
\begin{equation}
\label{eq:qhat}
\widehat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}.
\end{equation}
where $\textbf{R}$ is a weighting matrix chosen by the data analyst.  For example, inverse covariance weighting (the optimal form of QE) would set $\textbf{R} \equiv \textbf{C}^{-1}$ and a uniform-weighted case would use $\textbf{R} \equiv \textbf{I}$, the identity matrix.
The matrix $\textbf{Q}^{\alpha}$ encodes the dependence of the covariance on the power spectrum parameters $\partial \C/\partial p_\alpha$ but in practice also does other things, including implementing a transform of the frequency-domain visibilities to $\mathbf{k}$-space, taking into account cosmological scalings, and converting the visibilities from Jy to Kelvin.
%with dimensions ($N_{f}$,$N_{f}$) --- as

%%% COME BACK HERE
% Need to still explain the motivation for this
With an appropriate normalization 
%We normalize our power spectrum estimates using the matrix $\textbf{M}$:
\begin{equation}
\label{eq:phat}
\widehat{\textbf{P}} = \textbf{M}\widehat{\textbf{q}},
\end{equation}
$\widehat{\textbf{P}}$ is the estimate of the true power spectrum $\textbf{P}$. The data analyst has a choice for $
\textbf{M}$. For simplicity in this section we choose $\textbf{M}$ to be diagonal, although we explore other cases for the analysis 
of PAPER-64 data as explained in Section \ref{sec:Bias}, subject to the constraint that each element of $\widehat{\textbf{P}}$ is a weighted sum of the elements of $\textbf{P}$ with weights that sum to unity.
% JEA - is one really free to choose M and R independently?
